### 2024-10-10 data-visualization.qmd
20241010 - update season from 2023 to 2024

```{r}
pbp_regularSeason <- nfl_pbp %>% 
  dplyr::filter(
    season == 2023,
    season_type == "REG") %>%
  dplyr::filter(!is.na(posteam) & (rush == 1 | pass == 1))

@@

Defensive EPA per play during the 2023 NFL season is depicted as a function of offensive EPA per play in @fig-plotDefOffEPA.

```{r}
#| label: fig-plotDefOffEPA
#| fig-cap: "2023 NFL Offensive and Defensive EPA Per Play."

ggplot2::ggplot(
  data = epa_combined,
  aes(
    x = off_epa,
    y = def_epa)) +
  nflplotR::geom_mean_lines(
    aes(
      x0 = off_epa ,
      y0 = def_epa)) +
  nflplotR::geom_nfl_logos(
    aes(
      team_abbr = team),
      width = 0.065,
      alpha = 0.7) +
  labs(
    x = "Offense EPA/play",
    y = "Defense EPA/play",
    title = "2023 NFL Offensive and Defensive EPA per Play"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) + # horizontal y-axis title
  scale_y_reverse()
```

>>>>

```{r}
pbp_regularSeason <- nfl_pbp %>% 
  dplyr::filter(
    season == 2024,
    season_type == "REG") %>%
  dplyr::filter(!is.na(posteam) & (rush == 1 | pass == 1))

@@

Defensive EPA per play during the 2024 NFL season is depicted as a function of offensive EPA per play in @fig-plotDefOffEPA.

```{r}
#| label: fig-plotDefOffEPA
#| fig-cap: "2024 NFL Offensive and Defensive EPA Per Play."

ggplot2::ggplot(
  data = epa_combined,
  aes(
    x = off_epa,
    y = def_epa)) +
  nflplotR::geom_mean_lines(
    aes(
      x0 = off_epa ,
      y0 = def_epa)) +
  nflplotR::geom_nfl_logos(
    aes(
      team_abbr = team),
      width = 0.065,
      alpha = 0.7) +
  labs(
    x = "Offense EPA/play",
    y = "Defense EPA/play",
    title = "2024 NFL Offensive and Defensive EPA per Play"
  ) +
  theme_classic() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) + # horizontal y-axis title
  scale_y_reverse()
```

### 2024-10-10 sports-cognitive-psychology.qmd
20241010 - update season from 2023 to 2024

```{r}
# labels on the plot
text_df <- tibble(
  label = c("NFL coaches<br>in <span style='color:#00BFC4'>**2023**</span>", "NFL coaches<br>in <span style='color:#F8766D'>**2014**</span>"),
  x = c(6, 8.2),
  y = c(80, 37),
  angle = c(10, 10),
  color = c("black", "black")
)

nfl_4thdown %>%
  filter(vegas_wp > .2, between(go_boost, -10, 10), season %in% c(2014, 2023)) %>%
  ggplot(aes(go_boost, go, color = as.factor(season))) + 
  ggtext::geom_richtext(
    data = text_df,   
    aes(x, y, label = label, angle = angle), 
    color = "black",
    fill = NA,
    label.color = NA,
    size = 5) + 
  geom_vline(xintercept = 0) +
  stat_smooth(method = "gam", method.args = list(gamma = 1), formula = y ~ s(x, bs = "cs", k = 10), show.legend = F, se = F, linewidth = 4) +
  # this is just to get the plot to draw the full 0 to 100 range
  geom_hline(yintercept = 100, alpha = 0) +
  geom_hline(yintercept = 0, alpha = 0) +
  ggplot2::theme_classic() +
  labs(x = "Gain in win probability by going for it",
       y = "Go-for-it percentage",
       subtitle = "4th down decisions in 2023 versus 2014, win prob. > 20%",
       title = glue::glue("How <span style='color:red'>math</span> is changing football")) +
  theme(
    legend.position = "none",
    plot.title = element_markdown(size = 22, hjust = 0.5),
    plot.subtitle = element_markdown(size = 14, hjust = 0.5),
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold")
  ) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 4), limits = c(0, 100), expand = c(0,0)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10), limits = c(-10, 10), expand = c(0,0)) +
  annotate("text", x= -1.2, y= 70, label = "Should\nkick", color="black", size = 5) +
  annotate("text", x= 1.2, y= 70, label = "Should\ngo for it", color="black", size = 5) +
  geom_segment(
    aes(x = -.1, y = 80, xend = -2, yend = 80),
    arrow = arrow(length = unit(0.05, "npc")),
    color = "black", linewidth = 2
    ) +
  geom_segment(
    aes(x = .1, y = 80, xend = 2, yend = 80),
    arrow = arrow(length = unit(0.05, "npc")),
    color = "black", linewidth = 2
  )
```

>>>>

```{r}
# labels on the plot
text_df <- tibble(
  label = c("NFL coaches<br>in <span style='color:#00BFC4'>**2024**</span>", "NFL coaches<br>in <span style='color:#F8766D'>**2014**</span>"),
  x = c(6, 8.2),
  y = c(80, 37),
  angle = c(10, 10),
  color = c("black", "black")
)

nfl_4thdown %>%
  filter(vegas_wp > .2, between(go_boost, -10, 10), season %in% c(2014, 2024)) %>%
  ggplot(aes(go_boost, go, color = as.factor(season))) + 
  ggtext::geom_richtext(
    data = text_df,   
    aes(x, y, label = label, angle = angle), 
    color = "black",
    fill = NA,
    label.color = NA,
    size = 5) + 
  geom_vline(xintercept = 0) +
  stat_smooth(method = "gam", method.args = list(gamma = 1), formula = y ~ s(x, bs = "cs", k = 10), show.legend = F, se = F, linewidth = 4) +
  # this is just to get the plot to draw the full 0 to 100 range
  geom_hline(yintercept = 100, alpha = 0) +
  geom_hline(yintercept = 0, alpha = 0) +
  ggplot2::theme_classic() +
  labs(x = "Gain in win probability by going for it",
       y = "Go-for-it percentage",
       subtitle = "4th down decisions in 2024 versus 2014, win prob. > 20%",
       title = glue::glue("How <span style='color:red'>math</span> is changing football")) +
  theme(
    legend.position = "none",
    plot.title = element_markdown(size = 22, hjust = 0.5),
    plot.subtitle = element_markdown(size = 14, hjust = 0.5),
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold")
  ) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 4), limits = c(0, 100), expand = c(0,0)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10), limits = c(-10, 10), expand = c(0,0)) +
  annotate("text", x= -1.2, y= 70, label = "Should\nkick", color="black", size = 5) +
  annotate("text", x= 1.2, y= 70, label = "Should\ngo for it", color="black", size = 5) +
  geom_segment(
    aes(x = -.1, y = 80, xend = -2, yend = 80),
    arrow = arrow(length = unit(0.05, "npc")),
    color = "black", linewidth = 2
    ) +
  geom_segment(
    aes(x = .1, y = 80, xend = 2, yend = 80),
    arrow = arrow(length = unit(0.05, "npc")),
    color = "black", linewidth = 2
  )
```
